Chronic conditions and increased antibiotic consumption

### 2017-03-03 chronic-conditions/chronic-condition-drivers.Rmd
chronic conditions, better

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, 
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
```

@@

```{r summarize_users}
usage_summary = usage %>%
  filter(age >= 65) %>%
  group_by(bene_id) %>%
  summarize(n_claims=sum(n_claims), age=max(age)) %>%
  left_join(chronic_conditions, by='bene_id')
```

@@

  x %>%
    glm(n_claims ~ has_condition + age, family='poisson', data=.) %>%
    tidy %>%
    mutate(condition=condition)
}

@@

@@

```{r show_poisson}
res %>%
  filter(term=='has_conditionTRUE') %>%
  mutate(pct.increase=round((exp(estimate)-1.0)*100)) %>%
  select(condition, pct.increase, estimate, p.value) %>%
  arrange(desc(estimate)) %>%
  kable
```

@@

>>>>

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, 
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
library(psych)
library(gplots)
```

@@

```{r summarize_users}
usage_summary = usage %>%
  filter(age >= 65) %>%
  group_by(bene_id) %>%
  summarize(n_claims=sum(n_claims), age=max(age), sex=head(sex, 1)) %>%
  mutate(is_female=if_else(sex=='female', 1L, 0L)) %>%
  select(-sex) %>%
  left_join(chronic_conditions, by='bene_id')
```

```{r compute_correlations}
tetra = chronic_conditions %>%
  select(-bene_id) %>%
  mutate_all(function(x) x == 3) %>%
  tetrachoric()
```

@@

  x %>%
    glm(n_claims ~ has_condition, family='poisson', data=.) %>%
    tidy %>%
    mutate(condition=condition)
}

@@

```{r multivariate_model}
multivariate_formula = paste(c('n_claims ~ age + is_female', conditions), collapse=' + ') %>%
  formula

multivariate_res = usage_summary %>%
  lm(multivariate_formula, family='poisson', data=.) %>%
  tidy
```

## Correlations

Presumably many of the conditions are correlated. I computed tetrachoric correlations, which is a method designed to deal with correlations of dichotomous variables.

```{r, cache=FALSE}
tetra$rho %>% heatmap
```

Thankfully, Alzheimer's correlates strongly with Alzheimer's and dementia, and prostatic hyperplasia and prostate cancer correlate strongly negatively with breast cancer and endometrial cancer. Unfortunately, the negative correlations make everything else harder to see, so here's the same thing but without the male conditions.

```{r fig.width=7, fig.height=7, cache=FALSE}
(!(rownames(tetra$rho) %in% c('CNCRPRST', 'HYPERP'))) %>%
  { tetra$rho[., .] } %>%
  heatmap.2(trace='none')
```

COPD correlates strongly with asthma and lung cancer, but lung cancer and asthma do not correlate strongly with one another. There is a block of correlations focused around hypertension that includes ischemic heart disease, heart failure, anemia, chronic kidney disease, diabetes, hypertension, hyperlipidemia, and, importantly, COPD.

## Regression

### Univariate

First, I performed univariate regressions. In each case, the explanatory variable was the presence of the chronic condition.

@@

```{r show_poisson, cache=FALSE}
res %>%
  filter(term=='has_conditionTRUE') %>%
  mutate(pct.increase=round((exp(estimate)-1.0)*100)) %>%
  select(condition, pct.increase, estimate, p.value) %>%
  arrange(desc(estimate)) %>%
  kable
```

@@

### Multivariate

I also performed a multivariate Poisson regression, keeping all the chronic conditions as well as age and sex (coded as "is female") as explanatory variables.

Not surprisingly, COPD tops the list. However, asthma does provide explanatory power beyond COPD, coming in second. (Lung cancer, however, now has a negative effect on consumption.) Strangely, depression and anemia are next. Age, kindney disease, and diabetes have coefficients very near zero!

```{r cache=FALSE}
multivariate_res %>%
  mutate(pct.increase=round((exp(estimate)-1.0)*100)) %>%
  arrange(desc(pct.increase)) %>%
  kable
```
